\chapter{Approach}
In order to research the problem formulation of this paper, a practical approach has been used. To evaluate and compare the two different methods (as described in problem formulation) a mobile application with the same requirements has been developed in each method. Both of the mobile applications make use of the same web application. 
  
\section{The mobile applications}
The web application that is used composes of a single page and is functional for the web browser Google Chrome. The page displays a form with a text field, an image field and a location field. To get the image or the location value for the form the user can only press a button. When the image or location button is pressed the web application uses the Google Chrome’s built in ability to take a picture or get the location.

The mobile applications consists of two parts. The web application described above and the mobile application encapsulating the web application. The mobile application must encapsulate the web application and display the web application exactly the same as it is displayed in a browser. The mobile application must use the mobile’s native functions to get the value for the image and the location.

The text field  in the web application form must function the same way as it functions in Google Chrome. 

For the image the mobile application must give the user the option to use an existing image from the image gallery or take a picture using the camera on the phone. The image that is passed back to the web application must be of in Base64 format. 

The location must be obtained using the mobile’s GPS. When the user interacts with the mobile application the interaction must be directly with the web application encapsulated within the mobile application. 

In the project the following was used for developing and testing:
\begin{itemize}
\item Android SDK version 21
\item PhoneGap version 2.9.1
\item Apache Cordova CLI 4.0
\item Google Chrome 44.0.2403.157
\item Nexus 5
\item Android version: 5.1.1
\item Kernel version: 3.4.0-gbebb36b
\end{itemize}

\section{Measuring method}
The resulting mobile applications are measured using the software metric logical lines of code. The count that is used is:
% here we have a bullet list of what's included and what's excluded

The development methods are written in two different programming languages in the mobile application layer. Natively in Android is in Java and PhoneGap is in JavaScript. To compare the measurement result of logical lines of code we use the conversion table described by Galorath and Evans\cite[p.~163]{galorath2006}. JavaScript and Java is both third generation languages and therefore no conversion is needed. 

The logical lines of code written in the web application layer is also measured in both methods. To give an correct result we measure the added logical lines of code that has been written as a result of the mobile application. 

The resulting lines in the mobile and web application layer is added to give a total result for the developing method. To analyse the result of lines code is combined with our personal developing experience. 

\section{Communication and master slave relationship}
There was two important aspects that is focused on in the research. The master slave relationship between the web and mobile application layer. Where the web application is the master and mobile application layer is the slave. And handling the commun

To research 
An architecture for the mobile applications is suggested. The architecture allows the master and slave 


% old segment 
\section{Measuring method}
??
In order to compare the two development methods (described in problem formulation), we have tried using them both to 

In order to research the problem formulation of this paper, a practical approach has been used. The two different development methods (as described in problem formulation) have been used in a practical development process. The aim of the development has been to develop an architecture allowing independent web application code and mobile-application code to be written.

Apart from just developing architectures using the different methods, a demo-application was developed. The application had the same function requirements for both development methods.

This also put requirements on the web application, it had to confirm to a certain standard for communication with the mobile application in order for both applications to be able to use it. 

Another important aspect of the method, and a major part of the results found in this paper is the process of researching questions arising during the process. Examples of this can be questions regarding strengths and weaknesses of the method at hand, but also questions of a more basic nature, such as the behavior of function calls between the layers (web and mobile), are they synchronous?

\subsection{Demo application}
The demo application consists two mobile applications developed using the different development methods. 

The web application which is used in the mobile applicaitons is a simple form with a text field, an image field and a location field.

When running on a computer in a web browser the web application make use of the computer's camera and the web browser's own function to get the computers location.

The two mobile applications make use of the web application described above. To get the image for the web application form the mobile application makes use of the mobile's native camera function or uploads an image from the mobile's memory storage. To get the location the mobile applications makes use of the mobile's native GPS function. 

%Should this be in the result??
Application source code can be found at:
\begin{itemize}
  \item Web application: \url{https://github.com/Albin-trialbee/pollux-server}
  \item PhoneGap application: \url{https://github.com/DavidNorrestam/pollux-phonegap}
  \item Native Android mobile application: \url{https://github.com/buren-trialbee/pollux/}
\end{itemize}